Supplemental Materials

### 2021-07-20 supplement/supplement-main.Rmd
tweaks to plots

---
title: "Supplemental Materials"
subtitle: "Untitled vaccination intuitive theory paper"
author: "Derek Powell, Kara Weisman, and Ellen M. Markman"
date: "Jul 17, 2021 (DRAFT)"
output:
  pdf_document: 
    latex_engine: xelatex
    fig_caption: yes
bibliography: references.bib
csl: pnas.csl
figsintext : yes
always_allow_html: true
header-includes:
 \usepackage{float}
---

@@

library(DiagrammeR)
library(DiagrammeRsvg)

@@

# z$nodes_df
z2$nodes_df <- z2$nodes_df %>% mutate(label = recode_nodes(label))
z2$nodes_df$fontname <- "Helvetica"
z2$nodes_df$shape <- "ellipse"
z2$nodes_df$fillcolor <- ifelse(z2$nodes_df$label=="Vaccination\nIntentions", "grey","white")
z2$nodes_df$fixedsize <- FALSE
z2$edges_df$penwidth <- abs(edge_df_mod$strength)*2
z2$edges_df$penwidth <- 1.5
# z$edges_df$label <- round(edge_df_mod$value,2)
# z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
z2$edges_df$color <- map_to_brewer(edge_df_mod$strength, bins=9, option="RdPu")
# z2$edges_df$color <- map_to_viridis(edge_df_mod$strength, option="magma")
# 
# render_graph(z2)

@@

```{r disp-bootgraph, echo=F, fig.width=8, fig.height=5, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across 200 bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown."}

@@

Though we found the uncovered network structure to be quite revealing and informative, there are also some important limitations that should be acknowledged. In particular, the edges from the vaccine danger and vaccine effectiveness nodes to the vaccine toxicity node strike us as counterintuitive. Intuitively, we might imagine that vaccines having toxins could cause them to be dangerous and ineffective. Yet the edges in the model run from in the opposite direction, and so do not support a clear causal interpretation.

A Bayesian network is a factorized representation of a probability distribution and any given network is a member of a "Markov equivalence class"—a set of networks capable of factorizing the same distribution. This set of DAGs can be represented as a "completed partially directed acyclic graph" (CPDAG), a graph with both directed and undirected edges. The equivalence class for our model can be represented as a CPDAG with 14 directed edges and 15 undirected edges—indicating that the data do not constrain the directions of 15 of the edges in the final model. However, 10 of these edges’ directions are directly constrained by our top-down "generative model" constraints, which in turn constrains all but two of the remaining undirected edges, including the counterintuitive edges running toward vaccine toxins. This is also reflected in a bootstrapping analysis applying our chosen structure learning algorithm against 1000 bootstrapped replicates of our data. We found that the direction of some edges within the graph differed across bootstrapped runs (Holistic balance -- naturalism, medical skepticism -- parental expertise), but that these unintuitive relationships surrounding vaccine toxins were robust.

@@

h <- HydeNetwork(as.formula(bnlearn_to_hyde_string(main_model)))

@@

```{r main_model_plot, include=F}

z <- custom_plot.HydeNetwork(h, attr_theme="tb")

@@

# z$nodes_df
z$nodes_df <- z$nodes_df %>% mutate(label = recode_nodes(label))
z$nodes_df$fontname <- "Helvetica"
z$nodes_df$shape <- "ellipse"
z$nodes_df$fillcolor <- ifelse(z$nodes_df$label=="Vaccination\nIntentions", "grey","white")
z$nodes_df$fixedsize <- FALSE
z$edges_df$penwidth <- abs(edge_df_mod$value)*2
z$edges_df$penwidth <- 1.5
# z$edges_df$label <- round(edge_df_mod$value,2)
# z$edges_df$label <- edge_df_mod$label
# z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
z$edges_df$color <- map_to_brewer(edge_df_mod$value, bins=11, option="RdBu")

@@

```{r disp-maingraph, echo=F, fig.width=8, fig.height=5, fig.align='center', fig.cap="Bayesian network model of intuitive theory surrounding vaccination decisions. Edge color indicates the magnitude of the generative (blue) or preventive (red) coeffient for the links."}

@@

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in Figure \@ref(fig:disp-maingraph)) with an additional node representing the evidence. For instance, to model the effects of the “disease risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node. To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

@@

fit_s2_vaccIntent <- brm(
  response ~ phase*condition + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s2_reg,
  chains = 6,
  cores = 6,
  control = list(adapt_delta=.99),
  seed = 46179,
  file = "../local/fit_s2_vaccIntent.rds"
)

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the disease risk intervention on responses to the _Vaccination Intentions_ scale items, using the “brms” package for R [@burkner2017]. We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the Disease Risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention responses is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The disease risk intervention increased participants’ vaccination intention ratings (see Table \@ref(tab:s2_model)). We consider this to be a full replication of Horne et al.’s [-@horne.etal2015] results.

To model the changes in the belief network, we estimate the “evidence ratio” using the method described in the modeling section above. The maximum likelihood estimate of the evidence ratio was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the disease severity node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio. Given that our cognitive model is parameterized via noisy-logical functions, the predicted influence of each belief is closely related to its first-order correlation.

Figure \@ref(fig:s2_plot) (right) shows the correlation between predicted and observed belief changes conditional on the disease risk intervention.

@@

```{r echo=F, warning=F, message=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiment 2. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions. Right, mean observed and model predicted change scores for the disease risk intervention condition. In both plots, error bars represent 95\\% confidence intervals."}
## combined plot for paper
library(patchwork)

s2_diff_plot <- s2_scored %>% 
  spread(phase, mean) %>% 
  drop_na(scale) %>% 
  mutate(post = rescale_beta(post, -3, 3), pre = rescale_beta(pre, -3, 3)) %>% 
  # group_by(scale) %>% 
  # mutate(post = post/sd(post), pre = pre/sd(pre)) %>% 
  mutate(diff = post-pre) %>% 
  mutate(
    scale = recode_nodes2(scale),
    condition = ifelse(condition=="noInterv", "Control", "Disease Risk")
    ) %>% 
  group_by(condition, scale) %>%
  summarize(M = mean(diff, na.rm=TRUE), se = sd(diff)/sqrt(n())) %>% 
  ggplot(aes(x = reorder(scale,M), y = M, ymin = M-se*2, ymax=M+se*2, color=condition, shape=condition)) +
  geom_pointrange(position=position_dodge(width=.25)) +
  geom_hline(yintercept=0, linetype="dashed", color="grey") +
  scale_color_manual(values=c("#0067b1","#c50024")) +
  scale_shape_manual(values=c(17,16)) +
  theme_bw(base_size=11) +
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    panel.grid = element_blank(),
    legend.position = c(.35,.9),
    legend.direction = "horizontal",
    aspect.ratio=.5
    ) +
  labs(title="A", y = "Mean difference", x = "Belief scale", color="Intervention", shape="Intervention")

@@

cor_val <- round(cor(obs_pred$predicted, obs_pred$observed),3)

@@

Procedures for studies 3a and 3b closely followed those of Study 2, but these studies tested the effects of a novel intervention aimed at dispelling concerns about toxic additives in vaccines. Full text of this intervention is available in Appendix C of these supplementary materials. In both studies, participants were randomly assigned to either a no-intervention control condition or to the novel vaccine toxins intervention.

@@

Due to the similarity of their designs, we combined data from studies 3a and 3b for our primary analysis. However, for full transparency we first describe analyses for these studies individually. As in Experiment 2, we conducted a mixed-effects Bayesian ordinal regression to assess the effects of the vaccine toxins intervention on participants’ _Vaccination Intentions_ scores. The vaccine toxins intervention improved participant’s vaccination intention scores relative to a no-intervention control condition. Tables below show the results of these regression analyses for Experiment 3a, 3b, and the two studies combined. In each analysis, the coefficient representing the interaction between phase and condition is positive and credibly differs from zero, indicating that the vaccine toxins intervention positively impacted vaccination intentions.

@@

# -- TODO: recode phase/condition 5/25/21, 11:07 AM
s3_reg <- s3_long %>% 
  filter(scale=="vaccIntent") %>% 
  mutate(
    phase = relevel(factor(phase), ref="pre"),
    condition = relevel(factor(condition), ref="noInterv"),
    response = resp
    )

fit_s3_vaccIntent <- brm(
  response ~ phase*condition + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s3_reg,
  chains = 6,
  cores = 6,
  control = list(adapt_delta=.95),
  seed = 69568,
  file = "../local/fit_s3_vaccIntent.rds"
)

@@

fit_s3a_vaccIntent <- brm(
  response ~ phase*condition + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s3_reg %>% filter(study=="A"),
  chains = 6,
  cores = 6,
  control = list(adapt_delta=.99),
  seed = 69568,
  file = "../local/fit_s3a_vaccIntent.rds"
)

fit_s3b_vaccIntent <- brm(
  response ~ phase*condition + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s3_reg %>% filter(study=="B"),
  chains = 6,
  cores = 6,
  control = list(adapt_delta=.99),
  seed = 69568,
  file = "../local/fit_s3b_vaccIntent.rds"
)

@@

To model the effect of the vaccine toxicity intervention, we followed the same procedure as in Study 2, but this time augmented the model with a node representing the vaccine toxicity intervention as a child of the vaccine toxins belief node. The maximum likelihood estimate of the evidence ratio for the vaccine toxicity intervention was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the disease severity node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio.

Figure \@ref(fig:s3_plot) (right) shows the correlation between predicted and observed belief changes conditional on the vaccine toxicity intervention.

```{r s3_plot, echo=F, message=F, warning=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiments 3a and 3b. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions. Right, mean observed and model predicted change scores for the toxins intervention condition. In both plots, error bars represent 95\\% confidence intervals."}
s3_diff_plot <- s3_scored %>% 
  # spread(phase, mean) %>% 
  drop_na(scale) %>% 
  mutate(post = rescale_beta(post, -3, 3), pre = rescale_beta(pre, -3, 3)) %>% 
  mutate(diff = post-pre) %>% 
  mutate(
    scale = recode_nodes2(scale),
    condition = ifelse(condition=="noInterv", "Control", "Vaccine Ingredients")
    ) %>% 
  group_by(condition, scale) %>%
  summarize(M = mean(diff, na.rm=TRUE), se = sd(diff)/sqrt(n())) %>% 
  ggplot(aes(x = reorder(scale,M), y = M, ymin = M-se*2, ymax=M+se*2, color=condition, shape=condition)) +
  geom_pointrange(position=position_dodge(width=.25)) +
  geom_hline(yintercept=0, linetype="dashed", color="grey") +
  scale_color_manual(values=c("#0067b1","#c50024")) +
  scale_shape_manual(values=c(17,16)) +
  theme_bw(base_size=11) +
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    panel.grid = element_blank(),
    legend.position = c(.6,.15),
    legend.direction = "horizontal",
    aspect.ratio=.5
    ) +
  labs(title="A", y = "Mean difference", x = "Belief scale", color="Intervention", shape="Intervention")

@@

@@

fit_s4_vaccIntent <- brm(
  response ~ phase*vacc_news + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s4_reg %>% filter(scale=="vaccIntent"),
  chains = 6,
  cores = 6,
  control = list(adapt_delta=.99),
  seed = 40725,
  file = "../local/fit_s4_vaccIntent.rds"
)

@@

Mirroring our analyses of interventions, we regressed responses to the _vaccine intentions_ scale items on predictors for phase (pre-outbreak in August and September 2017 vs. post-outbreak May 2019, dummy-coded), news exposure (dummy-coded with non-exposure as the baseline), and an interaction between phase and news exposure onto participants’ scores, using a Bayesian mixed-effects ordinal regression including random intercepts by subject and item. To assess the effects of recent news exposure on vaccine intentions, we examined the interaction between phase and news exposure. Participants’ vaccine intentions were increased by recent news exposure compared to their responses 18 months ago (see Table \@ref(tab:s4_table)).

>>>>

---
title: "Supplemental Materials"
subtitle: "Untitled vaccination intuitive theory paper"
author: "Derek Powell, Kara Weisman, and Ellen M. Markman"
date: "Jul 18, 2021 (DRAFT)"
output:
  pdf_document: 
    latex_engine: xelatex
    fig_caption: yes
bibliography: references.bib
csl: pnas.csl
figsintext : yes
always_allow_html: true
header-includes:
 \usepackage{float}
---

@@

library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)

@@

# z$nodes_df
z2$nodes_df <- z2$nodes_df %>% mutate(label = recode_nodes(label))
z2$nodes_df$fontname <- "Helvetica"
z2$nodes_df$shape <- "ellipse"
z2$nodes_df$fillcolor <- ifelse(z2$nodes_df$label=="Vaccination\nIntentions", 
                                "#fffaba","white")
z2$nodes_df$fillalpha <- 0.1
z2$nodes_df$fixedsize <- FALSE
z2$edges_df$penwidth <- abs(edge_df_mod$strength)*2
z2$edges_df$penwidth <- 1.5
# z$edges_df$label <- round(edge_df_mod$value,2)
# z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
z2$edges_df$color <- map_to_brewer(edge_df_mod$strength, bins=9, option="RdPu")
# z2$edges_df$color <- map_to_viridis(edge_df_mod$strength, option="magma")
# 
# render_graph(z2)

@@

```{r disp-bootgraph, echo=F, fig.width=8, fig.height=5, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across 200 bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown. Our primary outcome of interest, vaccination intentions, is highlighted in yellow."}

@@

Though we found the uncovered network structure to be quite revealing and informative, there are also some important limitations that should be acknowledged. In particular, the edges from the vaccine danger and vaccine effectiveness nodes to the _vaccine toxicity_ node strike us as counterintuitive. Intuitively, we might imagine that vaccines having toxins could cause them to be dangerous and ineffective. Yet the edges in the model run from in the opposite direction, and so do not support a clear causal interpretation.

A Bayesian network is a factorized representation of a probability distribution and any given network is a member of a "Markov equivalence class"—a set of networks capable of factorizing the same distribution. This set of DAGs can be represented as a "completed partially directed acyclic graph" (CPDAG), a graph with both directed and undirected edges. The equivalence class for our model can be represented as a CPDAG with 14 directed edges and 15 undirected edges—indicating that the data do not constrain the directions of 15 of the edges in the final model. However, 10 of these edges’ directions are directly constrained by our top-down "generative model" constraints, which in turn constrains all but two of the remaining undirected edges, including the counterintuitive edges running toward _vaccine toxicity_. This is also reflected in a bootstrapping analysis applying our chosen structure learning algorithm against 1000 bootstrapped replicates of our data. We found that the direction of some edges within the graph differed across bootstrapped runs (Holistic balance -- naturalism, medical skepticism -- parental expertise), but that these unintuitive relationships surrounding _vaccine toxicity_ were robust.

@@

```{r, include=F}
h <- HydeNetwork(as.formula(bnlearn_to_hyde_string(main_model)))

@@

```{r main_model_plot, include=F}
z <- custom_plot.HydeNetwork(h, attr_theme="tb")

@@

# z$nodes_df
z$nodes_df <- z$nodes_df %>% mutate(label = recode_nodes(label))
z$nodes_df$fontname <- "Helvetica"
z$nodes_df$shape <- "ellipse"
z$nodes_df$fillcolor <- ifelse(z$nodes_df$label=="Vaccination\nIntentions", "#fffaba","white")
z$nodes_df$fixedsize <- FALSE
z$edges_df$penwidth <- abs(edge_df_mod$value)*2
z$edges_df$penwidth <- 1.5
# z$edges_df$label <- round(edge_df_mod$value,2)
# z$edges_df$label <- edge_df_mod$label
# z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
z$edges_df$color <- map_to_brewer(edge_df_mod$value, bins=11, option="RdBu")

@@

```{r disp-maingraph, echo=F, fig.width=8, fig.height=5, fig.align='center', fig.cap="Bayesian network model of intuitive theory surrounding vaccination decisions. Edge color indicates the magnitude of the generative (blue) or preventive (red) coeffient for the links. Our primary outcome of interest, vaccination intentions, is highlighted in yellow."}

@@

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in Figure \@ref(fig:disp-maingraph)) with an additional node representing the evidence. For instance, to model the effects of the “Disease Risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node. To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

@@

fit_s2_vaccIntent <- brm(
  response ~ phase*condition + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s2_reg,
  chains = 6,
  cores = 6,
  # control = list(adapt_delta=.99),
  seed = 46179,
  file = "../local/fit_s2_vaccIntent.rds"
)

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the Disease Risk intervention on responses to the _Vaccination Intentions_ scale items, using the “brms” package for R [@burkner2017]. We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the Disease Risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention responses is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The Disease Risk intervention increased participants’ vaccination intention ratings (see Table \@ref(tab:s2_model)). We consider this to be a full replication of Horne et al.’s [-@horne.etal2015] results.

To model the changes in the belief network, we estimate the “evidence ratio” using the method described in the modeling section above. The maximum likelihood estimate of the evidence ratio was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the _disease severity_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio. Given that our cognitive model is parameterized via noisy-logical functions, the predicted influence of each belief is closely related to its first-order correlation.

Figure \@ref(fig:s2_plot) (right) shows the correlation between predicted and observed belief changes conditional on the Disease Risk intervention.

@@

```{r echo=F, warning=F, message=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiment 2. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Disease Risk intervention condition. In both plots, error bars represent 95\\% confidence intervals."}
## combined plot for paper
library(patchwork)

s2_scale_order <- s2_scored %>% 
  drop_na(scale) %>% 
  spread(phase, mean) %>%
  mutate(post = rescale_beta(post, -3, 3), pre = rescale_beta(pre, -3, 3)) %>% 
  mutate(diff = post-pre) %>% 
  mutate(scale = recode_nodes2(scale)) %>% 
  group_by(condition, scale) %>%
  summarize(M = mean(diff, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(scale) %>%
  summarize(diff_diff = mean(M, na.rm = T)) %>%
  ungroup() %>%
  arrange(desc(diff_diff)) %>% # descending because intervention is expected to increase disease risk beliefs
  mutate(order = 1:nrow(.))

s2_diff_plot <- s2_scored %>% 
  spread(phase, mean) %>% 
  drop_na(scale) %>% 
  mutate(post = rescale_beta(post, -3, 3), pre = rescale_beta(pre, -3, 3)) %>% 
  # group_by(scale) %>% 
  # mutate(post = post/sd(post), pre = pre/sd(pre)) %>% 
  mutate(diff = post-pre) %>% 
  mutate(
    scale = recode_nodes2(scale),
    condition = recode_factor(condition, 
                              "noInterv" = "Control",
                              .default = "Disease Risk")) %>% 
  group_by(condition, scale) %>%
  summarize(M = mean(diff, na.rm=TRUE), se = sd(diff)/sqrt(n())) %>% 
  ungroup() %>%
  left_join(s2_scale_order) %>%
  ggplot(aes(x = reorder(scale,order), y = M, 
             ymin = M-se*2, ymax=M+se*2, 
             color=condition, shape=condition)) +
  scale_x_discrete() +
  annotate("rect", 
           xmin = s2_scale_order %>%
             filter(grepl("intent", tolower(scale))) %>% pull(order) - 0.25,
           xmax = s2_scale_order %>%
             filter(grepl("intent", tolower(scale))) %>% pull(order) + 0.25,
           ymin = -Inf,
           ymax = Inf,
           fill = "#fffaba") +
  geom_pointrange(position=position_dodge(width=.25)) +
  geom_hline(yintercept=0, linetype="dashed", color="grey") +
  scale_color_manual(values=c("#0067b1","#c50024")) +
  scale_shape_manual(values=c(16,17)) +
  theme_bw(base_size=11) +
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    panel.grid = element_blank(),
    legend.position = c(.5,.9),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = NA),
    aspect.ratio=.5
    ) +
  labs(title="A", y = "Mean difference", x = "Belief scale", 
       color="Intervention", shape="Intervention")

@@

cor_val <- format(round(cor(obs_pred$predicted, obs_pred$observed),3), nsmall = 3)

@@

Procedures for studies 3a and 3b closely followed those of Study 2, but these studies tested the effects of a novel intervention aimed at dispelling concerns about toxic additives in vaccines. Full text of this intervention is available in Appendix C of these supplementary materials. In both studies, participants were randomly assigned to either a no-intervention control condition or to the novel Vaccine Ingredients intervention.

@@

Due to the similarity of their designs, we combined data from studies 3a and 3b for our primary analysis. However, for full transparency we first describe analyses for these studies individually. As in Experiment 2, we conducted a mixed-effects Bayesian ordinal regression to assess the effects of the Vaccine Ingredients intervention on participants’ _Vaccination Intentions_ scores. The Vaccine Ingredients intervention improved participant’s vaccination intention scores relative to a no-intervention control condition. Tables below show the results of these regression analyses for Experiment 3a, 3b, and the two studies combined. In each analysis, the coefficient representing the interaction between phase and condition is positive and credibly differs from zero, indicating that the Vaccine Ingredients intervention positively impacted vaccination intentions.

@@

# -- TODO: recode phase/condition 5/25/21, 11:07 AM
s3_reg <- s3_long %>% 
  filter(scale=="vaccIntent") %>% 
  mutate(
    study = relevel(factor(study), ref="A"),
    phase = relevel(factor(phase), ref="pre"),
    condition = relevel(factor(condition), ref="noInterv"),
    response = resp
    )

contrasts(s3_reg$study) <- contr.sum(2)

fit_s3_vaccIntent <- brm(
  response ~ phase*condition*study + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s3_reg,
  chains = 6,
  cores = 6,
  # control = list(adapt_delta=.99),
  seed = 69568,
  file = "../local/fit_s3_vaccIntent.rds"
)

@@

fit_s3a_vaccIntent <- brm(
  response ~ phase*condition + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s3_reg %>% filter(study=="A"),
  chains = 6,
  cores = 6,
  # control = list(adapt_delta=.99),
  seed = 69568,
  file = "../local/fit_s3a_vaccIntent.rds"
)

fit_s3b_vaccIntent <- brm(
  response ~ phase*condition + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s3_reg %>% filter(study=="B"),
  chains = 6,
  cores = 6,
  # control = list(adapt_delta=.99),
  seed = 69568,
  file = "../local/fit_s3b_vaccIntent.rds"
)

@@

To model the effect of the Vaccine Ingredients intervention, we followed the same procedure as in Study 2, but this time augmented the model with a node representing the Vaccine Ingredients intervention as a child of the _vaccine toxicity_ belief node. The maximum likelihood estimate of the evidence ratio for the Vaccine Ingredients intervention was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the _vaccine toxicity_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio.

Figure \@ref(fig:s3_plot) (right) shows the correlation between predicted and observed belief changes conditional on the Vaccine Ingredients intervention.

```{r s3_plot, echo=F, message=F, warning=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiments 3a and 3b. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Vaccine Ingredients intervention condition. In both plots, error bars represent 95\\% confidence intervals."}

s3_scale_order <- s3_scored %>% 
  drop_na(scale) %>% 
  mutate(post = rescale_beta(post, -3, 3), pre = rescale_beta(pre, -3, 3)) %>% 
  mutate(diff = post-pre) %>% 
  mutate(scale = recode_nodes2(scale)) %>% 
  group_by(condition, scale) %>%
  summarize(M = mean(diff, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(scale) %>%
  summarize(diff_diff = mean(M, na.rm = T)) %>%
  ungroup() %>%
  arrange(diff_diff) %>% # ascending because intervention is expected to decrease vaccine toxicity beliefs
  mutate(order = 1:nrow(.))

s3_diff_plot <- s3_scored %>% 
  # spread(phase, mean) %>% 
  drop_na(scale) %>% 
  mutate(post = rescale_beta(post, -3, 3), pre = rescale_beta(pre, -3, 3)) %>% 
  mutate(diff = post-pre) %>% 
  mutate(
    scale = recode_nodes2(scale),
    condition = recode_factor(condition, 
                              "noInterv" = "Control",
                              .default = "Vaccine Ingredients")) %>% 
  group_by(condition, scale) %>%
  summarize(M = mean(diff, na.rm=TRUE), se = sd(diff)/sqrt(n())) %>% 
  ungroup() %>%
  left_join(s3_scale_order) %>%
  ggplot(aes(x = reorder(scale,order), y = M, ymin = M-se*2, ymax=M+se*2, 
             color=condition, shape=condition)) +
  scale_x_discrete() +
  annotate("rect", 
           xmin = s3_scale_order %>%
             filter(grepl("intent", tolower(scale))) %>% pull(order) - 0.25,
           xmax = s3_scale_order %>%
             filter(grepl("intent", tolower(scale))) %>% pull(order) + 0.25,
           ymin = -Inf,
           ymax = Inf,
           fill = "#fffaba") +
  geom_pointrange(position=position_dodge(width=.25)) +
  geom_hline(yintercept=0, linetype="dashed", color="grey") +
  scale_color_manual(values=c("#0067b1","#c50024")) +
  scale_shape_manual(values=c(16,17)) +
  theme_bw(base_size=11) +
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    panel.grid = element_blank(),
    legend.position = c(.5,.1),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = NA),
    aspect.ratio=.5
    ) +
  labs(title="A", y = "Mean difference", x = "Belief scale", 
       color="Intervention", shape="Intervention")

@@

```{r s3_plot_bystudy, echo=F, message=F, warning=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiments 3a and 3b. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Vaccine Ingredients intervention condition. In both plots, error bars represent 95\\% confidence intervals."}

# see previous chunk for scale order

s3_diff_plot_bystudy <- s3_scored %>% 
  # spread(phase, mean) %>% 
  drop_na(scale) %>% 
  mutate(post = rescale_beta(post, -3, 3), pre = rescale_beta(pre, -3, 3)) %>% 
  mutate(diff = post-pre) %>% 
  mutate(
    scale = recode_nodes2(scale),
    condition = recode_factor(condition, 
                              "noInterv" = "Control",
                              .default = "Vaccine Ingredients")) %>% 
  left_join(s3_demo %>% distinct(workerId, study)) %>%
  mutate(study = recode_factor(study, "A" = "Expt. 3A", "B" = "Expt. 3B")) %>%
  group_by(study, condition, scale) %>%
  summarize(M = mean(diff, na.rm=TRUE), se = sd(diff)/sqrt(n())) %>% 
  ungroup() %>%
  left_join(s3_scale_order) %>%
  ggplot(aes(x = reorder(scale,order), y = M, ymin = M-se*2, ymax=M+se*2, 
             color = interaction(condition, study, sep = ": "), 
             shape = interaction(condition, study, sep = ": "))) +
  scale_x_discrete() +
  annotate("rect", 
           xmin = s3_scale_order %>%
             filter(grepl("intent", tolower(scale))) %>% pull(order) - 0.35,
           xmax = s3_scale_order %>%
             filter(grepl("intent", tolower(scale))) %>% pull(order) + 0.35,
           ymin = -Inf,
           ymax = Inf,
           fill = "#fffaba") +
  geom_pointrange(aes(group = interaction(condition, study)),
                  position=position_dodge(width=.4), fill = "white") +
  geom_hline(yintercept=0, linetype="dashed", color="grey") +
  scale_color_manual(values=rep(c("#0067b1","#c50024"), 2)) +
  scale_shape_manual(values=c(16, 17, 21, 24)) +
  theme_bw(base_size=11) +
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    panel.grid = element_blank(),
    legend.position = c(.5,.1),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = NA),
    aspect.ratio=.5
    ) +
  labs(title="A", y = "Mean difference", x = "Belief scale", 
       color="Intervention", shape="Intervention") +
  guides(shape = guide_legend(nrow = 2, byrow = T),
         color = guide_legend(nrow = 2, byrow = T))

obs_pred_A <- make_pred_df(s3_mdf %>%
                             filter(workerId %in% s3_demo$workerId[s3_demo$study == "A"]), 
                           base_model, "vaccTox")
obs_pred_B <- make_pred_df(s3_mdf %>%
                             filter(workerId %in% s3_demo$workerId[s3_demo$study == "B"]), 
                           base_model, "vaccTox")

cor_val_A <- format(round(cor(obs_pred_A$predicted, obs_pred_A$observed),3), nsmall = 3)
cor_val_B <- format(round(cor(obs_pred_B$predicted, obs_pred_B$observed),3), nsmall = 3)

s3_obs_pred_plot_bystudy <- bind_rows(obs_pred_A %>% mutate(study = "A"),
                                      obs_pred_B %>% mutate(study = "B")) %>%
  mutate(study = recode_factor(study, "A" = "Expt. 3A", "B" = "Expt. 3B")) %>%
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul,
             shape = study)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_errorbar(alpha=.5, width=0) +
  geom_point(size=2.5, fill = "white") +
  scale_shape_manual(values=c(16, 21)) +
  # coord_fixed(ratio=1) +
  coord_cartesian(xlim=c(-.12,.08), ylim=c(-.12,.08)) +
  annotate(geom = "text", x = .05, y = -.08, 
           label = bquote(paste("r"["A"]*" = ", .(cor_val_A)))) +
  annotate(geom = "text", x = .05, y = -.105, 
           label = bquote(paste("r"["B"]*" = ", .(cor_val_B)))) +
  theme_bw(base_size=11) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = c(.15,.85),
        legend.background = element_rect(fill = NA)) +
  theme(aspect.ratio = 1) +
  labs(title = "B", y = "Observed change", x="Predicted change", 
       shape = "Experiment")

s3_plt_bystudy <- s3_diff_plot_bystudy + s3_obs_pred_plot_bystudy + 
  plot_layout(ncol=2, widths = c(2,1))

s3_plt_bystudy
```

@@

fit_s4_vaccIntent <- brm(
  response ~ phase*vacc_news + (1|item) + (1|workerId),
  family = cumulative("logit"),
  data = s4_reg %>% filter(scale=="vaccIntent"),
  chains = 6,
  cores = 6,
  # control = list(adapt_delta=.99),
  seed = 40725,
  file = "../local/fit_s4_vaccIntent.rds"
)

@@

Mirroring our analyses of interventions, we regressed responses to the _vaccine intentions_ scale items on predictors for phase (pre-outbreak in August and September 2017 vs. post-outbreak May 2019, dummy-coded), news exposure (dummy-coded with non-exposure as the baseline), and an interaction between phase and news exposure onto participants’ scores, using a Bayesian mixed-effects ordinal regression including random intercepts by subject and item. To assess the effects of recent news exposure on _vaccine intentions_, we examined the interaction between phase and news exposure. Participants’ vaccine intentions were increased by recent news exposure compared to their responses 18 months ago (see Table \@ref(tab:s4_table)).

To model the effects of news exposure we followed the same procedure as in previous studies, but here augmented the model with a node representing news exposure as a child of the _vaccine intentions_ node. The maximum likelihood estimate of the evidence ratio for news exposure was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the _vaccine intentions_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio.
